---
title: "Spam Project Analysis"
author: "K. Bret Staudt Willet"
date: "5/16/2019"
output: 
    html_document:
        toc: true
        float_toc: true
---

# Get set up

This section loads the data and packages and starts to process the data, but doesn't calculate any statistics or create any results.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
usethis::use_git_ignore(c("*.csv", "*.rds"))
```

## Load packages

```{r, include=FALSE}
library(tidyverse)
library(rtweet)
library(lubridate)
#library(janitor)
```

## Load the data

Having completed the steps in the setup.Rmd file, you now have the dataset stored in your local repository and can load it as usual. 

```{r, include=FALSE}
edchat_full <- read.csv("edchat_full_df.csv", header=TRUE, colClasses='character')
```

# Review the full dataset

```{r, include=TRUE}
n_tweets_full <- edchat_full %>% pull(id_str) %>%
    unique() %>% 
    length()
n_tweeters_full <- edchat_full %>% pull(from_user_id_str) %>%
    unique() %>% 
    length()
paste("Number of unique tweets:", n_tweets_full); paste("Number of unique tweeters:", n_tweeters_full)
```

# De-spam the data

## Clean with rtweet

Our first step of de-spamming our dataset is to run the data through the `rtweet` R package, which queries the Twitter API to return the most complete set of tweet metadata available. See https://rtweet.info/ for details on `rtweet`.

```{r, include=FALSE}
edchat_rtweet <- read.csv("edchat_rtweet_df.csv", header=TRUE, colClasses='character')
```

```{r, include=TRUE, echo=FALSE}
n_tweets_rtweet <- edchat_rtweet %>% pull(status_id) %>%
    unique() %>% 
    length()
n_tweeters_rtweet <- edchat_rtweet %>% pull(user_id) %>%
    unique() %>% 
    length()
paste("Number of unique tweets after rtweet:", n_tweets_rtweet)
paste("Number of unique tweeters after rtweet:", n_tweeters_rtweet)
```

```{r, include=TRUE, echo=FALSE}
tweet_loss_rtweet <- n_tweets_full - n_tweets_rtweet
tweeter_loss_rtweet <- n_tweeters_full - n_tweeters_rtweet
paste("Tweets lost:", tweet_loss_rtweet, "(", round(100 * tweet_loss_rtweet / n_tweets_full, 2), "% )")
paste("Tweeters lost:", tweeter_loss_rtweet, "(", round(100 * tweeter_loss_rtweet / n_tweeters_full, 2), "% )")
```

```{r, include=TRUE, echo=FALSE}
edchat_rtweet_public <- edchat_rtweet %>% filter(protected=="FALSE")
n_tweets_rtweet_public <- edchat_rtweet_public %>% pull(status_id) %>%
    unique() %>% 
    length()
n_tweeters_rtweet_public <- edchat_rtweet_public %>% pull(user_id) %>%
    unique() %>% 
    length()
paste("Number of public unique tweets after rtweet:", n_tweets_rtweet_public)
paste("Number of public unique tweeters after rtweet:", n_tweeters_rtweet_public)
```

```{r, include=TRUE, echo=FALSE}
tweet_loss_rtweet_public <- n_tweets_rtweet - n_tweets_rtweet_public
tweeter_loss_rtweet_public <- n_tweeters_rtweet - n_tweeters_rtweet_public
tweet_loss_rtweet_public; tweeter_loss_rtweet_public
paste("Tweets lost:", tweet_loss_rtweet_public, "(", round(100 * tweet_loss_rtweet_public / n_tweets_rtweet, 2), "% )")
paste("Tweeters lost:", tweeter_loss_rtweet_public, "(", round(100 * tweeter_loss_rtweet_public / n_tweeters_rtweet, 2), "% )")
```

## Rearrange data and calculate needed measures

```{r, include=FALSE, eval=FALSE}
freq_tweeters <- edchat_rtweet_public %>% 
    pull(user_id) %>% 
    table() %>% 
    as.data.frame() %>% 
    arrange(desc(Freq)) %>%
    rename(tweets_made_edchat = Freq)

pop_mean <- mean(freq_tweeters$tweets_made_edchat)

freq_tweeters <- freq_tweeters %>%
    mutate(sq_mean_diff = (tweets_made_edchat - pop_mean)^2)

pop_sd <- sqrt(mean(freq_tweeters$sq_mean_diff))

freq_tweeters <- freq_tweeters %>%
    mutate(z = (tweets_made_edchat - pop_mean) / pop_sd,
           edchat_prop = (tweets_made_edchat / n_tweets_rtweet_public) * 100
           ) %>%
    select(-sq_mean_diff)
freq_tweeters <- rename(freq_tweeters, user_id = .)
#hist(freq_tweeters$z)
#freq_tweeters %>% head(n=100)
```



```{r, include=FALSE}
# Define a function to not include NA values. 
length_with_na <- function(x) {
  ifelse(is.na(x), 0, map_int(x, length))
}
```



```{r, include=FALSE, eval=FALSE}
freq_replies <- edchat_rtweet_public %>% 
    pull(reply_to_status_id) %>% 
    table() %>% 
    as.data.frame() %>% 
    arrange(desc(Freq)) %>%
    rename(reply_count = Freq)
freq_replies <- rename(freq_replies, status_id = .)

edchat_rtweet_public <- edchat_rtweet_public %>% 
    full_join(freq_replies, by='status_id') %>% 
    mutate(reply_count = ifelse(is.na(reply_count), 0, reply_count))
```

```{r, include=FALSE, eval=FALSE}
edchat_tweeters <- edchat_rtweet_public %>% 
    mutate(tweets_made_all = statuses_count %>% as.numeric(),
           tweets_liked_all = favourites_count %>% as.numeric(),
           profile_description = description,
           favorite_count = ifelse(is.na(favorite_count), 0, favorite_count) %>% as.numeric(),
           retweet_count = ifelse(is.na(retweet_count), 0, retweet_count) %>% as.numeric(),
           hashtag_count = ifelse(is.na(hashtags), 0, strsplit(hashtags, " ") %>% sapply(length)), # length_with_na()
           hashtag_inclusion = ifelse(hashtag_count==0, 0, 1) %>% as.numeric(),
           url_count = ifelse(is.na(urls_url), 0, strsplit(urls_url, " ") %>% sapply(length)), # length_with_na()
           url_inclusion = ifelse(url_count==0, 0, 1),
           following_ratio = as.numeric(friends_count) / as.numeric(followers_count)
           ) %>%
    group_by(user_id) %>%
    mutate(like_mean = mean(favorite_count),
           retweet_mean = mean(retweet_count),
           reply_mean = mean(reply_count),
           hashtag_total = sum(hashtag_count),
           hashtag_mean = mean(hashtag_count),
           url_total = sum(url_count),
           url_mean = mean(url_count),
           tweets_with_hashtags = (sum(hashtag_inclusion) / n()) * 100,
           tweets_with_url = (sum(url_inclusion) / n()) * 100
              ) %>%
    slice(1) %>%
    full_join(freq_tweeters, by='user_id') %>%
    filter(!is.na(user_id)) %>%
    select(user_id, screen_name,
           tweets_made_all, tweets_made_edchat,
           z, edchat_prop,
           like_mean, retweet_mean, reply_mean,
           following_ratio, friends_count, followers_count,
           hashtag_total, hashtag_mean, tweets_with_hashtags, 
           url_total, url_mean, tweets_with_url,
           tweets_liked_all,
           source, verified,
           profile_description, profile_url
           ) %>%
    arrange(desc(z))

write.csv(edchat_tweeters, "edchat_tweeters.csv", row.names=FALSE)
```

```{r, include=FALSE}
edchat_tweeters <- read.csv("edchat_tweeters.csv", header=TRUE, colClasses= c(user_id='character'))
```

## Apply practical metrics for educational research

*Metric 1. Volume of tweeting*: One indicator of spam is unusually high-volume tweeting as such tweeting is often-bot generated. Related practical indicators of spam include counts of the raw number of tweets, the percentage of tweets to a hashtag accounted for by a user, or more standardized metrics such as z-scores of tweets per user.

```{r, include=TRUE}
# Total tweets by tweeter
#edchat_tweeters %>% arrange(desc(tweets_made_all)) %>% head(100) 

# Percentage of #Edchat volume by tweeter
#edchat_tweeters %>% arrange(desc(edchat_prop)) %>% head(100)

# Standard deviations away from the mean  (z-scores)
edchat_mean_tweets <- edchat_tweeters$tweets_made_edchat %>% mean() %>% round(2)
edchat_sd_tweets <- edchat_tweeters$tweets_made_edchat %>% sd() %>% round(2)
edchat_median_tweets <- edchat_tweeters$tweets_made_edchat %>% median()
paste("Mean:", edchat_mean_tweets); paste("SD:", edchat_sd_tweets); paste("Median:", edchat_median_tweets)

edchat_one_timers_n <- edchat_tweeters %>% filter(tweets_made_edchat==1) %>% nrow()
edchat_one_timers_p <- round(100 * edchat_one_timers_n / nrow(edchat_tweeters), 2)
paste("One-time tweeters:", edchat_one_timers_n, "(", edchat_one_timers_p, "% )")

edchat_sorted_by_z <- edchat_tweeters %>% arrange(desc(z))
#head(edchat_sorted_by_z, 100)
#View(edchat_sorted_by_z)
```

*Metric 2. Level of interaction*: Because spammers tend to broadcast messages, which others frequently ignore (Lin & Huang, 2013), spam accounts can also be identified by the absence of interaction with others. Relatively easy metrics researchers can use to measure interaction is to examine the extent to which a users’ tweets result in likes, retweets, and replies.

```{r, include=TRUE}
edchat_sorted_by_interaction <- edchat_tweeters %>% 
    arrange(desc(z), like_mean, retweet_mean, reply_mean) %>%
    select(user_id, screen_name, z, like_mean, retweet_mean, reply_mean)
#head(edchat_sorted_by_interaction, 100)
#View(edchat_sorted_by_interaction)

edchat_sorted_by_likes <- edchat_tweeters %>% 
    arrange(desc(like_mean), retweet_mean, reply_mean, z) %>%
    select(user_id, screen_name, z, like_mean, retweet_mean, reply_mean)
#head(edchat_sorted_by_likes, 100)
#View(edchat_sorted_by_likes)

edchat_sorted_by_retweets <- edchat_tweeters %>% 
    arrange(desc(retweet_mean), like_mean, reply_mean, z) %>%
    select(user_id, screen_name, z, like_mean, retweet_mean, reply_mean)
#head(edchat_sorted_by_retweets, 100)
#View(edchat_sorted_by_retweets)

edchat_sorted_by_replies <- edchat_tweeters %>% 
    arrange(desc(reply_mean), like_mean, retweet_mean, z) %>%
    select(user_id, screen_name, z, like_mean, retweet_mean, reply_mean)
#head(edchat_sorted_by_replies, 100)
#View(edchat_sorted_by_replies)

```

*Metric 3. Following vs. followers*: Spammers often follow many other users, but themselves have relatively low number of followers. Researchers can quickly measure this phenomenon by calculating the ratio of following to followers for users in their dataset.

```{r, include=TRUE}
edchat_no_followers_n <- edchat_tweeters %>% filter(following_ratio==Inf) %>% nrow()
edchat_no_followers_p <- round(100 * edchat_no_followers_n / nrow(edchat_tweeters), 2)
paste("Tweeters with no followers:", edchat_no_followers_n, "(", edchat_no_followers_p, "% )")

edchat_sorted_by_following_ratio <- edchat_tweeters %>% 
    arrange(desc(following_ratio), desc(z)) %>%
    select(user_id, screen_name, z, following_ratio, friends_count, followers_count, like_mean, retweet_mean, reply_mean)
#head(edchat_sorted_by_following_ratio, 100)
View(edchat_sorted_by_following_ratio)
```

*Metric 4. Level of hyperlinking*: Many spammers share hyperlinks in an attempt to drive traffic to certain websites (e.g., Lin & Huang, 2013) For instance, a tweet might advertise goods for sale and include a hyperlink to the website where the actual purchase would occur. Researchers can therefore analyze the raw number of links, the percentage of tweets that contain a link, or the average number of links per tweet.

```{r, include=TRUE}
edchat_sorted_by_hashtags <- edchat_tweeters %>% 
    arrange(desc(tweets_with_hashtags)) %>%
    select(user_id, screen_name, z, 
           tweets_made_all, tweets_made_edchat,
           hashtag_total, hashtag_mean, tweets_with_hashtags, 
           url_total, url_mean, tweets_with_url)
#head(edchat_sorted_by_hashtags, 100)
View(edchat_sorted_by_hashtags)

edchat_sorted_by_urls <- edchat_tweeters %>% 
    arrange(desc(url_mean)) %>%
    select(user_id, screen_name, z, 
           tweets_made_all, tweets_made_edchat,
           hashtag_total, hashtag_mean, tweets_with_hashtags, 
           url_total, url_mean, tweets_with_url)
#head(edchat_sorted_by_urls, 100)
View(edchat_sorted_by_urls)

```

## Remove the spammers

Explain why we chose these particular cutoffs...

```{r, include=TRUE}
edchat_closer_look <- edchat_tweeters %>% filter(z > 2, 
                                                 like_mean < 1, 
                                                 retweet_mean < 1,
                                                 reply_mean < 1,
                                                 following_ratio > 2
                                                 )
View(edchat_closer_look)
```

# Compare the full dataset with the de-spammed dataset

```{r, include=TRUE}
edchat_tweeters
```
